### 2024-05-26 README.Rmd
Testing textbox

---
output: github_document
bibliography: bibliography/bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{css textbox}
.blackbox {
  padding: 1em;
  background: black;
  color: white;
  border: 2px solid orange;
  border-radius: 10px;
}
.center {
  text-align: center;
}
```

```{r include=FALSE}
library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(magick) # Advanced Graphics and Image-Processing in R
library(neuralnet) # Training of Neural Networks
```

# Use of AI in Teaching and Learning: Some Considerations

<!-- badges: start -->

<!-- badges: end -->

By: Antonio Páez (paezha@mcmaster.ca)

Disclaimer: _All opinions expressed here are mine and any resemblance to those of my employer(s) is purely coincidental_.

## Background: An AI is Haunting Campus^[Karl Marx wrote about a "spectre...haunting Europe--the spectre of communism"; GenAI looks awfully like the shade of the spectre of unregulated surveillance capitalism.]

> Mission Statement: Our organization seeks to level up on all matters related.  
> -- An Organization Run by its Marketing Department

:::: {.blackbox data-latex=""}
::: {.center data-latex=""}
**NOTICE!**
:::

Thank you for noticing this **new notice**! Your noticing it has
been noted, and _will be reported to the authorities_!
::::

On November 30, 2022, OpenAI released what became its flagship product, [ChatGPT](https://en.wikipedia.org/wiki/ChatGPT), a chatbot based on [Large Language Models](https://en.wikipedia.org/wiki/Large_language_model) (LLMs) designed and trained to mimic natural language. The launch of the chatbot was an impressive feat of marketing [@cowen_chatgpt_2023], and it took less than two weeks for the hype to reach stratospheric levels: by early December, ChatGPT was being panned as "the best artificial intelligence chatbot ever released to the general public" [@roose_brilliance_2023], and a tool able to generate "impressively detailed" and "human-like" text, leading some to wonder whether it could "replace humans" [@lock_what_2022].

Universities, much like every other organization dealing with information, had to respond in a compressed span of time to the rapid release of generative AI (GenAI) tools like ChatGPT, most of them created and unleashed in the wild by a handful of large American corporations including [Meta](https://en.wikipedia.org/wiki/Meta_Platforms) (or were we to deadname it, facebook), [OpenAI](https://en.wikipedia.org/wiki/OpenAI), [Google](https://en.wikipedia.org/wiki/Alphabet_Inc.), and [Microsoft](https://en.wikipedia.org/wiki/Microsoft).

GenAI was announced as nothing short of revolutionary. McMaster, for example, [states](https://provost.mcmaster.ca/office-of-the-provost-2/generative-artificial-intelligence-2/generative-ai-for-staff/provisional-guidelines-on-the-use-of-generative-ai-in-operational-excellence/):

> Understanding how generative AI works, and how it can support work within organizations represents an ongoing area of exploration and innovation. The opportunities of generative AI are exciting: creating workplace efficiencies to allow for different kinds of work, adding capabilities for individuals and teams, and offering personalized uses.

AI was quickly perceived as an _innovative_ and _exciting_ technology, that moreover can lead to _efficiencies_^["Efficiencies", to be clear, is the day's euphemism for cutting costs [@oconnell_let_2012, p. 64].].  It seemed inevitable that universities _had to do something_ or risk being let behind.

The fear of missing out, combined with a modicum of concern for the ethical implications of GenAIs^[Ethical concerns for the most part have had to do with policing academic dishonesty among students], has led to institutional responses that have been less than coherent. An example is the case of Emory University, where the university gave a $10,000 prize to a group of students who proposed developing an AI tool for an Entrepreneurship Summit, only for the students to be subsequently suspended by the school's Honor Council for the potential of their tool for cheating^[There was no evidence that cheating had happened.] [@koebler__university_2024].

For its part, McMaster was a leader in responding to the rapid, and disruptive, release of these _revolutionary_ tools, and its provisional [guidelines](https://provost.mcmaster.ca/office-of-the-provost-2/generative-artificial-intelligence-2/generative-ai-for-staff/provisional-guidelines-on-the-use-of-generative-ai-in-operational-excellence/) (released under a Creative Commons Licence) have since inspired those of other post-secondary institutions in Canada, including the University of Guelph [-@office_of_teaching_and_learning_provisional_2023] and Trent University [-@trent_university_generative_2024].

The aim of the provisional guidelines was to inform the university community about the use of AI in "operational excellence". But the time has come to update the guidelines, and the institution, via its Office of the Provost, [launched a committee](#ai-committee-announced) to explore the use of AI across campus. This committee is composed of three sub-committees that examine the use of AI in 1) teaching and learning, 2) research, and 3) operational excellence in "our work" (the Provost's words)(probably meaning the university's work; not sure if the Provost themself uses AI).

The sub-committees, in turn, convened panels to:

1. [Provide recommendations]({#ai-panel-tasks}) on updates to the [Provisional Guidelines](https://provost.mcmaster.ca/office-of-the-provost-2/generative-artificial-intelligence-2/generative-ai-for-staff/provisional-guidelines-on-the-use-of-generative-ai-in-operational-excellence/)

2. And also to do (what? consultation?).

3. Q: What are the terms of reference of the panels? Who will decide which recommendations are adopted and which are ignored?

According to the announcement of this AI Committee at McMaster:

> AI is a powerful tool and I believe that universities have an important role to play in its adoption.  
> -- Susan Tighe, Provost

Accepting that AIs are powerful tools, right off the bat, is consistent with the hype observed so far [@angwin_press_2024]. Hammers are powerful tools. Word processors are powerful tools. Lasers are powerful tools. And yet we do not have high-level committees for each of and every tool that the university uses. What makes AIs different from other tools that they require a pan-university consultation on an "_ethical approach to shaping policies and developing new ways of using AI to ensure a responsible and beneficial integration where it makes sense_"?

(The last sentence in the statement above carries quite a bit of weight here: where does adoption of AI make sense? Presumably this is what the Committee on AI will try to answer.)

<!--As I argue here, AI does not make a lot of sense outside of some limited applications that are nowhere near the scope of the promises of the merchants of hype.-->

To understand why many organizations, including McMaster University, have decided to spend considerable resources studying the adoption and responsible use of AI, it is essential to have a working understanding of what AI does that sets it apart from other tools.

In doing this, we must be alert to the fact that technologies are never politically neutral: it is important to ask two questions:

- What does a specific technology do for whom?

- And what does it do to whom?

We also need to understand the goals of the organization that considers the technology. When does it make sense for a university like McMaster to "integrate" AI as part of the core mission of the organization, accounting for what it does to some people and what it does for others?

Let us begin with the mission of McMaster as an institution of higher education.

## The [mission](https://president.mcmaster.ca/mission-vision/)

McMaster's mission is "the discovery, communication and preservation of knowledge," accomplished with a commitment to "creativity, innovation and excellence". These aspirations must respond to some underlying values, and indeed, the university explicitly values "integrity, quality, inclusiveness and teamwork". Ultimately, the university aims to "inspire critical thinking, personal growth, and a passion for lifelong learning" and to serve "the social, cultural, and economic needs of our community and our society."

A mission statement is like an onion, and one needs to peel multiple layers to try to understand the core principles enunciated there. Why are things like "integrity" and "critical thought" valued? Or, what happens when "integrity", "quality", and/or "excellence" clash with "the economic needs of our society" (say, a mandate by democratically elected governments to spend as little as possible in post-secondary education^[Between 2000-2020, Ontario had the lowest spending on post-secondary education of all provinces in Canada, relative to its share of spending [@caut_provincial]. Public funding of universities in Ontario as percent of revenue fell from 45.6% in 2001/2 to 35.5% in 2019/20 [@statistics_canada_college_2022]. It is little wonder that universities are seeking "efficiencies"])?

Clearly, there will be some tension implicit in the mission of a university. But being generous, let us take the mission statement at face value and assume that the highest priorities are those related to knowledge obtained, preserved, and communicated with integrity and quality.

## What do AIs do?

At this point, we might as well ask "what do AIs _not_ do"?

A possible reason why many organizations have decided to spend substantial resources considering the adoption of GenAI is that, unlike most other tools, they are seen as _everything tools_. Indeed, the release of GenAIs to the masses was accompanied by innumerable lofty promises, albeit most of them made by entities that had a stake in the adoption of the technology.

The public was promised that GenAI would "accelerate sustainability" [@nakagawa_accelerating_2023]; that LLMs would transform sectors as diverse as health care, finance, human resources, insurance, "and beyond" by "automating customer self-service, accelerating response times on an increasing number of tasks as well as providing greater accuracy, enhanced routing and intelligent context gathering" [@ibm_what_2023]; according to Allerin^[Curiously, Allerin is not identified as an artificial intelligence company in @joshi_5_2022 but only in the much later article by @garsten_what_2024.], AI-powered self-driving cars would reduce accidents, car ownership, pollution and noise, and would automate logistics and make parking tickets obsolete [@joshi_5_2022; @garsten_what_2024]. In mental health care, chatbots could complement human therapists "via comprehensive data access and [analysis of] behavioral patterns" to "mimic practitioner questions and subsequently make recommendations based on a user’s inputs" [@silva_4_2023].

The list of promises has grown to encompass pretty much everything, from saving the planet all the way to micromanaging intimate facets of individual lives, for example by revolutionizing dating through the use of chatbots to help "people connect and find potential partners" (who presumably also use chatbots to do the same) [@khalatian_matchmaking_2023].

Higher-education was not spared the onslaught of promises.

@power_3_2024,^[Rhett Power is a motivational speaker and executive coach] for example, presents three innovative use cases for AI in higher education. First, AI could take over manual tasks and improve the budget by releasing administrative staff^[Probably to work on something else, and not for the university, hence leading to "efficiencies"]. AI could evaluate applications for admissions, including assessing traits like the "grit" and "empathy" of applicants. AI could also personalize the experience of consumers (also called students) by predicting their final grade with 60-70% accuracy, and by providing corrective and instructive feedback.
<!--release them in a very literal sense, as one advantage is to do _more_ with limited resources (which in the current fiscal environment means laying people off). The cases discussed by Mr. Power are questionable: there are already non-AI systems that deploy paperwork; the kind of information needed to identify students at risk of dropping out is probably quite invasive, and therefore impractical or unethical to collect; I very much doubt that it is legal to use AI in hiring and admissions. --> Mr. Power's message to the higher-education sector? It is time for institutions to "level up": jump on the bandwagon, fellas, don't let it let you behind!

In another instance, @hie_how_2023 argue that AIs will likely be the primary way to access information in the future. For this reason, institutions must equip students with the skills needed to use AI competently, in particular to perfect the art of _prompt engineering_, that is, the ability to craft effective queries that prompt a GenAI to provide the most useful outputs. Universities will use AIs to improve learning and challenge the critical thinking of students. Faculty will also benefit from AI: it will be used to evaluate their students^[A task of dubious legality, and clearly a way to cheat students of a learning experience], and will handle repetitive tasks like writing course syllabi^[In my 20+ years of experience as an educator, writing syllabi has been essential to understand my own courses; automating the task is another way of cheapening the quality of the learning experience for students]. @hie_how_2023^[Anthony Hié is Chief Innovation and Digital officer at [Excelia](https://www.excelia-group.com/about-excelia/our-schools/excelia-business-school). Writing in 2022, Mr. Hié predicted that the metaverse would be a [revolution in higher education](https://www.linkedin.com/pulse/metaverse-dedicated-pedagogy-revolution-higher-education-anthony-hi%C3%A9). In October 2023 facebook [slashed its metaverse team](https://www.reuters.com/technology/meta-lay-off-employees-metaverse-silicon-unit-wednesday-2023-10-03/). Microsoft [disbanded](https://www.theinformation.com/articles/microsoft-kills-its-industrial-metaverse-team-after-4-months) its Industrial Metaverse Core Team four months after launching it. Claire Thouary is the founder of [QACE UP](http://qace-up.fr), a consulting firm that supports higher-education institutions in their quest for "quality excellence".] warn that "AI is here to stay", that "[faculty] might not be able to stop the AI trend from growing", and so they might as well get on with the program and try to shape it as best they can.

What does McMaster hope to achieve with AI? From the Overview on Generative AI document:

> Some of [the uses of AI] in our context could include qualitative and quantitative data analysis, writing text (e.g. reference letters, grant applications, job descriptions, report summaries), translation, captioning, multiple formats for learning and engagement, and personalized support.

## But really, what do AIs do?

GenAIs are _deep learning models_ [@ibm_what_2021], so called because they are neural networks with _depth_, meaning many, many hidden layers. As an example, Large Language Models (such as those behind OpenAI's ChatGPT, Microsoft's Copilot, Google's Gemini, and facebook's LLaMA) are models built with multiple layers of neural networks trained on _very_ large amounts of textual data [@ibm_what_2023].

But, what are [neural](https://en.wikipedia.org/wiki/Large_language_model#Further_reading) [networks](https://medium.com/@asifurrahmanaust/lesson-3-neural-network-is-nothing-but-a-linear-regression-e05a328a0f23)?

Behind all the jargon, the neural networks that underpin GenAI are models designed to produce "statistically probable outputs" when prompted [@ibm_what_2021]. In more conventional terms, neural networks are regression models [@ripley_neural_1994]--much more sophisticated and powerful than linear regression to be sure, but regression models nonetheless. Here, it is important to note that the term regression refers to [regression to the mean](https://en.wikipedia.org/wiki/Regression_toward_the_mean)--regressing a prompt to a "statistically probable" outcome.

A regression model takes data as inputs, typically a "dependent" variable $y$ (observations of an outcome of interest) and "independent" variables $x$ (observations of things thought to correlate with the dependent variable). With these inputs, the model finds a conditional mean for the output $\hat{y}$ (i.e., a mean value conditional on the values of the dependent variables). The process of finding that conditional mean is to satisfy some critarion (e.g., that the conditional mean minimizes the error of the model--the distance between $y$ and $\hat{y}$).

Below is an example of the simplest regression model with only one independent variable (i.e., a bivariate model). The plot shows the pairs of $y$ and $x$ values, and the line is the model, the conditional mean $\hat{y}$. This model needs only two parameters: an intercept (the value of &\hat{y}$ when $x = 0$), and the slope of the line, in other words, the rate of change of $\hat{y}$ with respect to $x$. Each additional independent variable uses one additional parameter to describe the slope of the model with respect to that variable.
```{r simulate-simple-example}
set.seed(13253)

b0 <- 1.0
b1 <- 1.5

df <- data.frame(x = b0 + b1 * runif(100, 
                                0, 
                                10),
                 e = rnorm(100),
                 type = "training data") |>
  mutate(y = x + e)

x <- df |> pull(x)
y <- df |> pull(y)

mod <- lm(y ~ x, 
          data = df)
```

```{r plot-simulated-data, message=FALSE, out.width="50%", fig.cap="Simple Example of a Regression Model"}
ggplot(df,
       aes(x = x,
           y = y)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  theme_minimal()
```

Predict:
```{r data-vs-predictions-comparison, out.width="50%", fig.cap="Training data and model output compared"}

df_pred <- data.frame(x = x)

y_hat <- predict(mod, df_pred)

df_pred_lr <- df_pred |> 
  mutate(y = y_hat,
         type = "regression output")

df_pred_all <- rbind(df |>
                   select(x, y, type),
                 df_pred_lr |>
                   select(x, y, type))

ggplot() + 
  geom_point(data = df_pred_all,
  aes(x = x,
      y = y,
      color = type,
      shape = type,
      size = type)) +
  scale_color_manual(values = c("training data" = "black",
                                "regression output" = "blue",
                                "nn1 output" = "chocolate1",
                                "nn2 output" = "brown")) +
  scale_shape_manual(values = c("training data" = 16,
                                "regression output" = 17,
                                "nn1 output" = 18,
                                "nn2 output" = 19)) +
  scale_size_manual(values = c("training data" = 1,
                                "regression output" = 2,
                                "nn1 output" = 2,
                                "nn2 output" = 2)) +
  theme_minimal()
```

Neural networks do something similar but using a net-like architecture that is more flexible at identifying the characteristics of the data. For instance, the following figure shows a relatively simple neural network trained using using the same data. This network is not "deep", as it has only one so called "hidden layer", comprised of two neurons between the input and the output layers. In comparison to the linear regression and its two parameters, this neural network uses seven parameters.

```{r}
# Given the dataframe grid.1, let us now train a neural network. We need to first define a formula that specifies the dependent variable and the independent variables.
f <- as.formula("y ~ x")
```

```{r train-shallow-neural-network}
# Train a neural network using only two neurons in one hidden layer. The argument `linear.output` is set to TRUE, so that the activation function is not applied to the output node. The default activation function is the logistic function.
set.seed(33225433)

nn1 <- neuralnet(f, 
                data = df, 
                hidden = c(2), 
                linear.output = T) 
```

```{r save-shallow-neural-network}
# For some reason, the figure cannot be saved directly, so I saved it manually using screen capture
plot(nn1)
```

```{r plot-shallow-neural-network, out.width="50%", fig.cap="Simple Example of a Shallow Neural Network"}
knitr::include_graphics("images/nn1.png")
```

```{r shallow-neural-network-predictions}
y_hat <- predict(nn1,
                 df_pred)

df_pred_nn1 <- df_pred |> 
  mutate(y = y_hat,
         type = "nn1 output")

df_pred_all <- rbind(df_pred_all |>
                   select(x, y, type),
                 df_pred_nn1 |>
                   select(x, y, type))
```

The plot below illustrates the predictions of this neural network, where we can see that it is not limited to a linear response; neural networks are more flexible than linear regression, and they achieve this by being less parsimonious (they use more parameters).

```{r data-vs-predictions-comparison-2, out.width="50%", fig.cap="Training data and model outputs compared: linear regression and shallow neural network"}
ggplot() + 
  geom_point(data = df_pred_all,
  aes(x = x,
      y = y,
      color = type,
      shape = type,
      size = type)) +
  scale_color_manual(values = c("training data" = "black",
                                "regression output" = "blue",
                                "nn1 output" = "chocolate1",
                                "nn2 output" = "brown")) +
  scale_shape_manual(values = c("training data" = 16,
                                "regression output" = 17,
                                "nn1 output" = 18,
                                "nn2 output" = 19)) +
  scale_size_manual(values = c("training data" = 1,
                                "regression output" = 2,
                                "nn1 output" = 2,
                                "nn2 output" = 2)) +
  theme_minimal()
```

More complex neural network can be designed. The following plot is of a deeper network than the previous one, with two hidden layers, each consisting of five neurons. This model uses forty-six parameters. 
```{r train-deeper-neural-network}
# Train a neural network using only two neurons in one hidden layer. The argument `linear.output` is set to TRUE, so that the activation function is not applied to the output node. The default activation function is the logistic function.
set.seed(242)
nn2 <- neuralnet(f, 
                data = df, 
                hidden = c(5, 5), 
                linear.output = T)
```

```{r save-plot-deeper-neural-network}
# For some reason, the figure cannot be saved directly, so I saved it manually using screen capture
plot(nn2)
```
```{r plot-deeper-neural-network, out.width="50%", fig.cap="Example of a Deeper Neural Network"}
knitr::include_graphics("images/nn2.png")
```

```{r}
y_hat <- predict(nn2,
                 df_pred)

df_pred_nn2 <- df_pred |> 
  mutate(y = y_hat,
         type = "nn2 output")

df_pred_all <- rbind(df_pred_all |>
                   select(x, y, type),
                 df_pred_nn2 |>
                   select(x, y, type))

```

The increased complexity leads to greater flexibility, and we can see that the model makes predictions that resemble more closely the actual data.

```{r data-vs-predictions-comparison-3, out.width="50%", fig.cap="Training data and model outputs compared: shallow neural network and deeper neural network"}
ggplot() + 
  geom_point(data = df_pred_all |>
               filter(type != "regression output"),
  aes(x = x,
      y = y,
      color = type,
      shape = type,
      size = type)) +
  scale_color_manual(values = c("training data" = "black",
                                "regression output" = "blue",
                                "nn1 output" = "chocolate1",
                                "nn2 output" = "brown")) +
  scale_shape_manual(values = c("training data" = 16,
                                "regression output" = 17,
                                "nn1 output" = 18,
                                "nn2 output" = 19)) +
  scale_size_manual(values = c("training data" = 1,
                                "regression output" = 2,
                                "nn1 output" = 2,
                                "nn2 output" = 2)) +
  theme_minimal()
```

```{r}
# How do these models perform?
mad_lr <- data.frame(df |>
             mutate(y_hat = df_pred_lr$y)) |>
  summarize(mse = sum(abs(y - y_hat)/n()))

mad_nn1 <- data.frame(df |>
             mutate(y_hat = df_pred_nn1$y)) |>
  summarize(mad = sum(abs(y - y_hat)/n()))

mad_nn2 <- data.frame(df |>
             mutate(y_hat = df_pred_nn2$y)) |>
  summarize(mad = sum(abs(y - y_hat)/n()))

delta_mad_nn1 <- (mad_nn1 - mad_lr)/mad_lr * 100

delta_mad_nn2 <- (mad_nn2 - mad_lr)/mad_lr * 100
```

Indeed, the first neural network is "better" in that it reduces the typical error by `r paste0(prettyNum(delta_mad_nn2, digits = 2), "%")` compared to the linear regression, whereas the deeper network reduces the typical error by `r paste0(prettyNum(delta_mad_nn2, digits = 2), "%")`, also with respect to the linear regression^[The "typical error" here is the mean absolute deviation of the predictions.].

There is a limit to how flexible any model can be, and this is as true of linear regression as it is of neural networks. Each additional parameter makes the model more flexible, but as a rule we cannot use more parameters than we have data to train the model. Roughly, every additional parameter needs a data input to train on, and when there are as many parameters as data inputs the model is overfit and will predict each value of the independent variable in the training dataset perfectly.

<!-- 
IBM uses a dataset called Project CodeNet Dataset, which is sourced from two online judge web sites, AIZU Online Judge and AtCoder. https://research.ibm.com/blog/codenet-ai-neurips-2021 According to IBM "[the] end goal of CodeNet is to enable developers to create systems that can modernize existing codebases, as well as fix errors and security vulnerabilities in code."

From Aizu Terms of Use: https://onlinejudge.u-aizu.ac.jp/term_of_use

Handling of Submitted and Registered Codes

Commercial uses of the registered codes are prohibited.

You acknowledge and agree that users of this site may use copies or modifications of the public codes only for educational or research purposes.

You acknowledge and agree that public source codes may be opened collectively to persons authorized to access only for educational or research purposes.

You acknowledge and agree that source code submitted may be used by University of Aizu and related research institutes only for educational or research purposes.

IBM studiously avoids the word "commercial" in its posts about LLM, foundational models, and CodeNet. Is IBM doing all this research as a non-commercial effort? Will those developers creating systems, modernizing databases, fixing errors, etc., be able to access this for free? What does "commercial uses" even mean?
-->

**NOTE**: Labeling data is an essential, yet extremely expensive aspect of training GenAIs.

## What do GenAI tools do for students at McMaster?

## What do GenAI tools do to students at McMaster?

## Not learning

Words go here.

## Learning the wrong stuff

Words go here.

## Deskilling

Words go here.

### Persistence of skills (how long will learning to use a chatbot be useful?)

How useful will be AIs in the future? Think of model poisoning, and the possibility that we are going to see (or may already have seen) Peak ChatGPT.

De-duplication of training data improves LLMs [@lee2022deduplicating]. No wonder: it amounts to having more different information.

But there is the following implication: once human-generated content is exhausted and/or has been overwhelmed by chatbot generated content, de-duplication becomes increasingly challenging.

The variance of the model's output is substantially reduced: the output will, of necessity, be less varied, less diverse, and more norm-conforming:
```{r data-vs-prediction-variance}
df_pred_all |> 
  group_by(type) |> 
  summarize(variance = var(y))
```

## What do GenAIs do for the mega-corporations pushing these tools?

Words go here.

## What do GenAIs to to everyone else?

### Extractive

In relation to the points above (GenAIs require vast amounts of _human-generated_ data to ensure that there is sufficient variability for the model to train) - (Models trained )

@germain_google_2023
@norem_intels_2023

### Exploitative

Above, the ways in which GenAI is exploitative of people all over the world, but particularly in the Global South were highlighted. But AI companies do not spare the rich either. OpenAI's Sam Altman, a self-avowed admirer of the movie _Her_, featuring Scarlet Johansson as an AI voice assistant, reportedly tried to hire Ms. Johannson to be the voice of ChatGPT 4.0 [@kastrenakes_scarlett_2024]. After Ms. Johannson decline, OpenAI went and used a mimicry of [her voice](@as-an-aside) anyways, which was retired after Ms. Johansson threatened legal action^[Altman's denials that OpenAI had not used Ms. Johansson's voice sounded hollow, after he [posted](https://x.com/sama/status/1790075827666796666) on social media the word "her" at the time the new version of ChatGPT was announced.] [@allyn_scarlett_2024]. Alas, very few have the social standing of Scarlett Johansson to challenge OpenAI. Most people in the world lack that power.

@williams_exploited_2022 (Summary needed)

@perrigo_2_2023 (Follows summary)

ChatGPT was trained on vast amounts of data scooped from the internet, where many societal problems are amplified. Trained on horrible content, it was hard to sell ChatGPT when it was prone to racist, sexist, obscene outbursts. But companies like facebook, who did not want to pay for moderation, had hit on a solution of sorts <!--: train an AI to detect content that was unpalatable.-->

The solution?

> [F]eed an AI with labeled examples of violence, hate speech, and sexual abuse, and that tool could learn to detect those forms of toxicity in the wild. That detector would be built into ChatGPT to check whether it was echoing the toxicity of its training data, and filter it out before it ever reached the user. It could also help scrub toxic text from the training datasets of future AI models.

The key here is "labeled examples". What are those labels? Many machine algorithms do not learn by themselves...they learn from humans, and need what is called "labelled" data: a images that humans have labelled as "dog", "cats", etc. Or bits of text that have been labelled as...well...questionable. Labelling data is extremely time consuming and labor intensive. What's a corporation to do?

> To get those labels, OpenAI sent tens of thousands of snippets of text to an outsourcing firm in Kenya, beginning in November 2021. Much of that text appeared to have been pulled from the darkest recesses of the internet. Some of it described situations in graphic detail like child sexual abuse, bestiality, murder, suicide, torture, self harm, and incest. "

What did this mean to the people doing all this labeling in one of the poorest regions in the world? First of all, an income: "data labelers employed...on behalf of OpenAI were paid a take-home wage of between around $1.32 and $2 per hour depending on seniority and performance." But at what cost?

> One Sama worker tasked with reading and labeling text for OpenAI told TIME he suffered from recurring visions after reading a graphic description of a man having sex with a dog in the presence of a young child. “That was torture,” he said. “You will read a number of statements like that all through the week. By the time it gets to Friday, you are disturbed from thinking through that picture."

@hao_hidden_2023

Quotes from Hao's podcast:

> Karen Hao: One of the reasons why ChatGPT was able to become so virally popular and continued to sustain popularity is because it is largely not spewing really awful things. People feel comfortable using the product knowing that it's not going to do that.

> Annie Minoff: At least it won't do it in English. If Alex wanted to use ChatGPT in his native language, Swahili, would he be able to do that?

> Karen Hao: You can interact with ChatGPT in Swahili, but ChatGPT was developed primarily to work in English. So a lot of the scrubbing, the content moderation, the important safety measures within the chatbot were done in English. So when you prompt it in Swahili, you'll get more misinformation. You'll get more confusing sentences that don't make sense, and you will potentially get more of this content that they worked so hard to filter out because they were only filtering it in English.

Think about linguistic hegemony, and how the models are only as good (or as bad) as their inputs.

### Copyright (and privacy) issues

LLM are known to regurgitate verbatim chunks of the training datasets. This is called _memorization_, a word that contributes to antropomorphize what is essentially a statistical model. Memorization, in fact, is just a fancy way of saying "model overfitting"--in other words, an extreme regression to the mean that excels at predicting data in the training set, but is poor at projecting new data points. Overfitting happens when a model has too many parameters or too few data inputs (Models have [billions of parameters](https://en.wikipedia.org/wiki/Large_language_model#Training_cost) and are trained on [billions of pages of text](https://www.ibm.com/topics/large-language-models])). Data memorization--the verbatim or quasi-verbatim output of training data--can be as high as 7% in the case of some LLMs [@peng_near_duplicate_2023, p. 8].

Where will new data come from? [@murgia_why_2023]

Note how several organizations have responded to the AIfication of all the things: by banning it. The concerns cited are seldom ethical or moral, but mostly about the risk of sensitive information leaking. As LLMs begin to be increasingly trained on model-produce data, overfitting (or memorization) can become a larger risk.

And this is just a result of using an overfitted model. What about data leaks? Some stories regarding Microsoft's security track record:

https://www.theregister.com/2024/04/12/microsoft_cisa_order/

https://techreport.com/news/microsoft-azure-hit-with-the-largest-data-breach-in-its-history-hundreds-of-executive-accounts-compromised/

Also, how confident can the public be about assurances that a big tech corporation will not data mine and commercialize data by the companies that sell the services of LLMs? Microsoft is so big that it can stand up to the US government:

https://www.theregister.com/2024/04/05/microsoft_government_contracts/

<!-- "Nice operation you have there...be a shame if your systems stopped working all of a sudden."-->

Organization ranging from Free Open Source projects to elements of the federal government in the United States have limited/restricted the use of AI.

NetBSD (a project that produces a free, Unix-like Open Source operating system) states in its commit guidelines (contributions to code) that "[c]ode generated by a large language model or similar technology...is presumed to be tainted code, and must not be committed without prior written approval by core." [@netbsd_netbsd_nodate].

The National Archives and Records Administration (USA) bans use of ChatGPT [@koebler_national_2024]

US Congress bans use of ChatGPT and Copilot for staff [@singh_after_2024]. This article also note how companies like Samsung and Apple have restricted the use of chatbots citing concerns with data security.

### Environmental impacts

Microsoft: Here's how burning the planet will help to save it.

https://blogs.microsoft.com/on-the-issues/2024/05/15/microsoft-environmental-sustainability-report-2024/ [@smith_our_2024]

https://blogs.microsoft.com/on-the-issues/2023/11/16/accelerating-sustainability-ai-playbook/ [@nakagawa_accelerating_2023]

"Environmental costs of AI are soaring" [@crawford_generative_2024]

ChatGPT consumes substantially more energy than a regular web search, up to 25 times more, and according to some estimates, the whole of AI may consume twice as much energy as the whole of France by 2030 [@brussels_chatgpt_2024].

...are the results better?

Compare the results:

```{r puzzle-web-search, out.width="50%", fig.cap="\\label{fig:puzzle-web-search}A search using Duck Duck Go"}
ggdraw() + 
  draw_image(image = "images/logical-puzzle-web-search.png")
```

```{r puzzle-chatgpt, out.width="50%", fig.cap="\\label{fig:puzzle-chatgpt}A conversation with ChatGPT-4"}
puzzle_plot <- ggplot() +
  labs(caption = "https://assemblag.es/@hugo/112450401657084750") +
  theme_void()

ggdraw() + 
  draw_plot(puzzle_plot) +
  draw_image(image = "images/logical-puzzle-chat-gpt-4.png")

```

The example above may seem like an amusement, but a GenAI is not a toy, as Air Canada discover when the company was found liable for "negligent misrepresentation", after a court denied claims by the airline that the chatbot was somehow responsible for its own actions as an effectively separate legal entity [@garcia_what_2024].

Similar GenAI-generated mishaps abound.

But what when an AI replicates the both-sides perspective of much of the media when trying to learn about an existential crisis like climate changes? [@gordon_multiple_2024]. Or when trying to obtain accurate voting information in the context of voter supression? [@angwin_seeking_2024]

### Replication of unjust structures of power

When using a Generative AI we must ask whose perspectives that tool represents.

## As an aside {#as-an-aside}

On antropomorphizing GenAI tools. This happens mostly in ways that are barely liminal. For example, in the document "Overview of AI at McMaster" we read that GenAI's somehow have an "ability to engage in nuanced conversation". This phrasing could be simply a literary device, but even so it is indicative of a tendency to ascribe human attributes to a statistical model. Doing this risks amplifying existing biases. But in other cases, the effort is more blatant. Wrapping a statistical model in a human-like skin may actually be more devious, as in the new, sexy ChatGPT interface:

<video width="320" height="240" controls>
<source src="video/open-ai-commercial.mp4" type="video/mp4">
</video>

**NOTES:** Need to research more the argument that the interface is so important in the case of these tools because what they do well is trivial and what they do poorly is what is being hyped. Found the article [@farrugia_known_2024]; see here: https://fasterandworse.com/known-purpose-and-trusted-potential/

## Concluding remarks

ChatGPT was an awesome feat of marketing [@cowen_chatgpt_2023], but much coverage of the relevant issues in the media has been so credulous it has stretched credulity past Pluto [@burneko_if_2024]^[Bob Woodward said "journalism is not stenography" [@hiskes_woodward_2007], but it is difficult to see what else writing like Roose's [-@roose_is_2024] could be called.]. There are now calls to press pause on the hype [@angwin_press_2024] but these are still drown by the credulous coverage.

I suggest that universities should avoid the mistakes of the media, an industry that has innovated itself [out of the trust of the public](https://reutersinstitute.politics.ox.ac.uk/news/julia-angwin-fears-public-sphere-about-get-worse-ai-makes-it-easier-flood-zone-misinformation) (also see the "pivot-to-video" and how tech companies suckered many media companies, including ).

Why? The reputational risk...universities are under siege from unsympathetic or actively hostile actors. Universities main source of power is their reputation as places that preserve and expand knowledge in a principled way. "This university uses ChatGPT" is a way to say that the university is redundant...an organization as expensive as an institution of higher-education is not needed when there are cheaper alternatives for such an education ("act now before you are laid off!"):

```{r ligemy-ai-bundle, out.width="50%", fig.cap="\\label{fig:ai-success-bundle}An AI success bundle from the \"The Real People of Ligency PR and Marketing Team\""}
bundle_plot <- ggplot() +
  labs(caption = "https://www.udemy.com/user/ligency-team/") +
  theme_void()

ggdraw() + 
  draw_plot(bundle_plot) +
  draw_image(image = "images/trail-blazing-ai-bundle.png")
```

"McMaster is best known for our pioneering work in problem-based learning, which began in our medical school in 1969 and has been adopted across the world." https://provost.mcmaster.ca/teaching-learning/

Do we really want to be best known in 40 years as "McMaster, the institution that pioneered the use of ChatGPT"?

Recall that [regression to the mean](https://en.wikipedia.org/wiki/Regression_toward_the_mean) is also called "reversion to mediocrity" [@barnett_regression_2017].

## Loose thoughts

Are GenAI reality-warping machines? Think about the small and large ways in which these tools can be used to rewrite history, for relatively banal commercial objectives (monetizing clicks) or worse, for malicious propaganda campaigns.

**NOTE** Russia copied millions of articles from wikipedia wholesale, to replace the open encyclopedia with a home-grown version without all the bits that the current regime finds objectionable [@corfield_russia_2023; @jankowicz_russia_2023].

>The first step in liquidating a people,' said Hubl, 'is to erase its memory. Destroy its books, its culture, its history. Then have somebody write new books, manufacture a new culture, invent a new history. Before long the nation will begin to forget what it is and what it was. The world around it will forget even faster.  
>-― Milan Kundera, The Book of Laughter and Forgetting

Potential for weaponization of GenAI. **Farfetched (?) idea** Could it be that GenAI will be at some point declared a military-grade technology with restricted applications? (see https://www.defenseone.com/technology/2024/05/ai-and-other-tech-creating-special-operations-renaissance/396388/)

## Notes

### AI Committee announcement {#ai-committee-announced}

This is the announcement of the launch of the AI Committee (see email from the Office of the Provost, dated March 20, 2024):

<!--https://app.fastmail.com/mail/search:expert+panels/T2d9b9367655ae218.M6205d28844daa67dfd3432ff-->

> "McMaster launches AI committee
>
> Whether you are a researcher, faculty member, student, administrator, or have another role on campus, you are likely thinking about Artificial Intelligence (AI) and its challenges and potential.
>
> AI is a powerful tool and I believe that universities have an important role to play in its adoption. At McMaster, we are well placed to prioritize an ethical approach to shaping policies and developing new ways of using AI to ensure a responsible and beneficial integration where it makes sense.
>
> I’m pleased to announce that McMaster has launched an AI Advisory Committee to explore the use of AI across campus. Three subcommittees have also been started to examine the use of AI in teaching and learning, research and operational excellence in our work. It is a strategy that makes McMaster one of the few institutions taking a pan-university approach to AI.
>
> Our earlier work to develop provisional guidelines for the use of generative AI in teaching and learning drew interest from other post-secondary institutions, which adopted McMaster's guidelines after we made them available through a Creative Commons licence.
>
> The committee will be led by three co-chairs, each bringing different expertise. They are Matheus Grasselli, deputy provost, Gayleen Gray, associate vice-president and chief technology officer and Gianni Parise, acting deputy vice-president, Research. They will be supported by Erin Aspenlieder, special advisor to the provost on generative AI.
>
> The three subcommittees will each by led by a subject matter expert. Kim Dej, vice-provost, Teaching and Learning, will lead the exploration of AI in teaching and learning, Maggie Pooran, executive director, Health, Safety, Well-Being & Labour Relations, will explore its use in operational excellence, and Martin Horn, associate dean, Faculty of Humanities, will look at implications for research.
>
> Each expert subcommittee will have a pool of contributors made up of members of the McMaster community. Depending on the initiative, some or all members will be invited to complete the specific work based on their expertise, interest and availability.
>
> If you would like to be considered for the expert panels, please complete this expression of interest form."

### Panel tasks{#ai-panel-tasks}

The task of the Artificial Intelligence Advisory Committee Expert Panel on Teaching and Learning was initially defined as follows by the Vice-Provost Teaching and Learning in an email dated April 16:

<!--https://app.fastmail.com/mail/search:erin/T6fae26856703ef40.Mfe5dff646c6ae452b84ef4bf-->

> [T]he first task of this Expert Panel will be to provide recommendations to the AI Advisory Committee on updates to the Provisional Guidelines on the Use of Generative AI in Teaching and Learning.

<!--This does not say anything about what the Committee will do with the recommendations; I cannot find anything about this.-->

## References