### 2024-07-02 paper/highlights.qmd
Paper submitted to Computers & Education

---
title: "An AI is Haunting Campus: A Call for Human-Centered Higher Education"
subtitle: "Highlights"
format: docx
---

- AI has been presented as an _everything tool_ in higher education
- AIs are regression models: they use inputs to regress to a probable output, i.e., a mean 
- AIs cannot achieve the variety of tasks that are promised without compromising the values of universities
- Students are poorly served by all the attention paid to AIs
- AIs are poor complements to a human-centered education